A GRASP algorithm for the concrete delivery problem.

### 2023-08-05 Article/article.tex
save CQ

\documentclass{article}
% \documentclass[review]{elsarticle}
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage[algo2e,noend]{algorithm2e}
\usepackage[compatible,noend]{algpseudocode}
\usepackage{enumitem}
\usepackage{xspace}
\usepackage[table,xcdraw]{xcolor}
\usepackage{amsfonts}
\usepackage{pifont}
\usepackage{bbding}
\usepackage[english]{babel}
\usepackage{adjustbox}
\usepackage{tikz}
\usetikzlibrary{automata, topaths, calc, positioning, shapes, backgrounds, fit, matrix}
\usepackage{caption}
\usepackage{tabularx}
\usepackage[T1]{fontenc}
\usepackage{natbib}
\usepackage{datatool}
\usepackage{array}
\usepackage{pgfplotstable}
\pgfplotsset{width=7cm,compat=newest}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{afterpage}
\usepackage{capt-of}
\usepackage{multirow}
\usepackage{float}
\usepackage[text={15cm,21cm},centering]{geometry}
\usepackage{setspace}
\usepackage{array}
\allowdisplaybreaks
\usepackage{lineno}
\usepackage{tkz-graph}
\usepackage{breqn}
\tikzset{
base/.style = {shape=rectangle, 
                      anchor=center,minimum size=5mm, 
                     align=center,
                     top color=white, ,inner sep=0ex,
                     minimum size=5mm
                     },
    LC11/.style =   {  base,   
                    text width=3em, 
                    bottom color=brown!100,
                    },
    UC11/.style = {     base,    
                    text width=5em, 
                    bottom color=brown!100,
                    align = right,
                },
    LC12/.style =   {  base,   
                    text width=1em, 
                    bottom color=orange!100,
                    },
    UC12/.style = {     base,    
                    text width=2.5em, 
                    bottom color=orange!100,
                },
    LC21/.style =   {  base,   
                    text width=1.5em, 
                    bottom color=brown!100,
                    },
    UC21/.style = {     base,    
                    text width=3.5em, 
                    bottom color=brown!100,
                },
    LC22/.style =   {  base,   
                    text width=4em, 
                    bottom color=blue!100,
                    },
    UC22/.style = {     base,    
                    text width=7em, 
                    bottom color=blue!100,
                    align = left,
                },
    UC221/.style = {     base,    
                    text width=3.5em, 
                    bottom color=blue!100,
                    % color=blue!100,
                    outer xsep=0em,
                    % align = right,
                },
                UC231/.style =   {  base,   
                    text width=1.5em, 
                    bottom color=blue!100,
                    },
    LC23/.style =   {  base,   
                    text width=1.5em, 
                    bottom color=orange!100,
                    },
    UC23/.style = {     base,    
                    text width=2.em, 
                    bottom color=orange!100,
                },
  env/.style      = {base, font=\ttfamily\normalsize},
label/.style   = {base,font=\ttfamily\normalsize, text width = 8em, 
  inner sep=0em, rounded corners=1mm,bottom color=red!80,},
  dummy/.style    = {circle,draw}
}

@@

Regarding objectives, most authors have focused on minimizing travel time and delays between consecutive deliveries. However, some authors have been more interested in maximizing customer satisfaction alone. We find these situations in the works of \cite{durbin2008or, kinable2014concrete, kinable2014logic, sulaman2017simulated}. \cite{kinable2014concrete} introduce a general MIP and constraint programming (CP) models of the CDP reflecting the main constraints commonly found in all CDP works: time lag and no overlapping between consecutive deliveries, covering of all customers' demands, delivery time window, and heterogeneous fleet. However, the model did not include constraints limiting the time that concrete may reside in a truck. The authors propose a constructive heuristic that schedules the visits to the customers one by one according to the start time of the visit and the truck capacity. The procedure is invoked multiple times for different permutations of the customer's order which is determined using the steepest descent (SD) local search procedure. One of the paper's main contributions is the creation of the first public test instances for the CDP with up to 50 customers, four batch plants, and 20 concrete mixers. They found the CP model to be highly effective in finding high-quality solutions in a relatively short time or improving existing schedules, while the MIP model can be used to compute bounds, as it seems ineffective in solving large problem instances. Finally, the heuristic often yields good solutions in less than a second. A detailed analysis of the MIP model presented in \cite{kinable2014concrete} and of two more compact models can be found in the thesis of \cite{hernandez_lopez_study_2020}. In \cite{kinable2014logic}, we found an attempt to solve the previous problem with a logic-based Benders' approach. \cite{sulaman2017simulated} expand upon the SD heuristic proposed in \cite{kinable2014concrete}, proposing a simulated annealing (SA) combined with a time-slot Heuristic (SATH). This method looks for a slot between existing visits of a truck to schedule a new delivery instead of assigning it to the time slot strictly after the truck's latest assigned delivery. The goal is to reduce the large time gaps that can be present in a schedule created with SD due to ignoring the intermediate available time slots. Experimental results indicated that SATH outperforms SD in speed and solution quality. A generalization of the MIP model of \cite{kinable2014concrete} is addressed in \cite{asbach2009analysis}. This model simultaneously minimizes the total sum of travel costs and the penalty costs for customers with unfulfilled demand. A customer can request that all concrete deliveries come from the same plant or a subset of plants and that a delivery truck belongs to a subset of the vehicle fleet. The MIP model is used in a local search scheme as a black-box solver to reoptimize an incumbent solution in which a neighborhood operator has unfixed some variables. \cite{tzanetos2023systematic} provide an overview of the various methods used in the literature to address the CDP and categorizes the problem formulations based on the different concepts used in the literature. They also discussed the consistency between industry needs and existing constraints and provided insights into the datasets corresponding to real-world cases, identifying the necessary data for practitioners.

@@

In this section, we present and discuss the results obtained in the computational experiments of the GRASP heuristic described in Section~\ref{grasp_method}. The GRASP algorithm was coded in C++. We run our experiments on dataset extracted from data provid by our industrial partner.
% and the branch-and-bound (B\&B) algorithm applied to the model described in Section~\ref{model}. The different algorithms were coded in C++ and run on a processor Intel $2.4$ GHz Gold $6148$ Skylake. For the exact method, we have used IBM Concert Technology and CPLEX $12.10$ as MIP solver. A single thread was used with a time limit of one hour.
% No benchmark instances for the \dirp exist so we have generated the dataset for the experiments as follows.

@@

In order to evaluate the effectiveness and performance of our GRASP algorithm, we obtained a dataset from our partner that represents the delivery operations of 8 production centers for a period of 42 days.

We generate 42 instances, each corresponding to a day operation. An instances has informations about the number of production centers, customers, orders, and drivers.
For each depot, we have its ID and loading capacity. For a customer, we have the ID of the assigned depot, his demand, due time and number of orders. Then for each order we have his demand. For the driver, we have the capacity and shift start time.

@@

@@

\cite{kinable2014concrete} provided a dataset for their variant of the $CDP$. as remainder, their objective is to maximize the total load served each day, and there is no time allocated for loading operations at a plant. While the original objective of the dataset may not align directly with our specific objective, we can still utilize it to analyze and compare the performance of our algorithm under different circumstances.

The results are below.

>>>>

\documentclass{article}
% \documentclass[review]{elsarticle}
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage[algo2e,noend]{algorithm2e}
\usepackage[compatible,noend]{algpseudocode}
\usepackage{enumitem}
\usepackage{xspace}
\usepackage[table,xcdraw]{xcolor}
\usepackage{amsfonts}
\usepackage{pifont}
\usepackage{bbding}
\usepackage[english]{babel}
\usepackage{adjustbox}
\usepackage{tikz}
\usetikzlibrary{automata, topaths, calc, positioning, shapes, backgrounds, fit, matrix}
\usepackage{caption}
\usepackage{tabularx}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{datatool}
\usepackage{array}
\usepackage{pgfplotstable}
\pgfplotsset{width=7cm,compat=newest}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{afterpage}
\usepackage{capt-of}
\usepackage{multirow}
\usepackage{float}
\usepackage[text={15cm,21cm},centering]{geometry}
\usepackage{setspace}
\usepackage{array}
\allowdisplaybreaks
\usepackage{lineno}
\usepackage{tkz-graph}
\usepackage{breqn}
\tikzset{
base/.style = {shape=rectangle, 
                      anchor=center,minimum size=5mm, 
                     align=center,
                     top color=white, ,inner sep=0ex,
                     minimum size=5mm
                     },
    LC11/.style =   {  base,   
                    text width=3em, 
                    bottom color=brown!100,
                    },
    UC11/.style = {     base,    
                    text width=5em, 
                    bottom color=brown!100,
                    align = right,
                },
    LC12/.style =   {  base,   
                    text width=1em, 
                    bottom color=orange!100,
                    },
    UC12/.style = {     base,    
                    text width=2.5em, 
                    bottom color=orange!100,
                },
    LC21/.style =   {  base,   
                    text width=1.5em, 
                    bottom color=brown!100,
                    },
    UC21/.style = {     base,    
                    text width=3.5em, 
                    bottom color=brown!100,
                },
    LC22/.style =   {  base,   
                    text width=4em, 
                    bottom color=blue!100,
                    },
    UC22/.style = {     base,    
                    text width=7em, 
                    bottom color=blue!100,
                    align = left,
                },
    UC221/.style = {     base,    
                    text width=3.5em, 
                    bottom color=blue!100,
                    % color=blue!100,
                    outer xsep=0em,
                    % align = right,
                },
                UC231/.style =   {  base,   
                    text width=1.5em, 
                    bottom color=blue!100,
                    },
    LC23/.style =   {  base,   
                    text width=1.5em, 
                    bottom color=orange!100,
                    },
    UC23/.style = {     base,    
                    text width=2.em, 
                    bottom color=orange!100,
                },
  env/.style      = {base, font=\ttfamily\normalsize},
label/.style   = {base,font=\ttfamily\normalsize, text width = 8em, 
  inner sep=0em, rounded corners=1mm,bottom color=red!80,},
  dummy/.style    = {circle,draw}
}

@@

Regarding objectives, most authors have focused on minimizing travel time and delays between consecutive deliveries. However, some authors have been more interested in maximizing customer satisfaction alone. We find these situations in the works of \cite{durbin2008or, kinable2014concrete, kinable2014logic, sulaman2017simulated}. \cite{kinable2014concrete} introduce a general MIP and constraint programming (CP) models of the CDP reflecting the main constraints commonly found in all CDP works: time lag and no overlapping between consecutive deliveries, covering of all customers' demands, delivery time window, and heterogeneous fleet. However, the model did not include constraints limiting the time that concrete may reside in a truck. The authors propose a constructive heuristic that schedules the visits to the customers one by one according to the start time of the visit and the truck capacity. The procedure is invoked multiple times for different permutations of the customer's order which is determined using the steepest descent (SD) local search procedure. One of the paper's main contributions is the creation of the first public test instances for the CDP with up to 50 customers, four batch plants, and 20 concrete mixers. They found the CP model to be highly effective in finding high-quality solutions in a relatively short time or improving existing schedules, while the MIP model can be used to compute bounds, as it seems ineffective in solving large problem instances. Finally, the heuristic often yields good solutions in less than a second. A detailed analysis of the MIP model presented in \cite{kinable2014concrete} and of two more compact models can be found in the thesis of \cite{hernandez_lopez_study_2020}. In \cite{kinable2014logic}, we found an attempt to solve the previous problem with a logic-based Benders' approach. \cite{sulaman2017simulated} expand upon the SD-heuristic proposed in \cite{kinable2014concrete}, proposing a simulated annealing (SA) combined with a time-slot Heuristic (SATH). This method looks for a slot between existing visits of a truck to schedule a new delivery instead of assigning it to the time slot strictly after the truck's latest assigned delivery. The goal is to reduce the large time gaps that can be present in a schedule created with SD due to ignoring the intermediate available time slots. Experimental results indicated that SATH outperforms SD in speed and solution quality. A generalization of the MIP model of \cite{kinable2014concrete} is addressed in \cite{asbach2009analysis}. This model simultaneously minimizes the total sum of travel costs and the penalty costs for customers with unfulfilled demand. A customer can request that all concrete deliveries come from the same plant or a subset of plants and that a delivery truck belongs to a subset of the vehicle fleet. The MIP model is used in a local search scheme as a black-box solver to reoptimize an incumbent solution in which a neighborhood operator has unfixed some variables. \cite{tzanetos2023systematic} provide an overview of the various methods used in the literature to address the CDP and categorizes the problem formulations based on the different concepts used in the literature. They also discussed the consistency between industry needs and existing constraints and provided insights into the datasets corresponding to real-world cases, identifying the necessary data for practitioners.

@@

In this section, we present and discuss the results of the computational experiments of the GRASP heuristic described in section~\ref{grasp_method}. The GRASP algorithm is coded in C++. We run our experiments on a dataset extracted from data provided by our industrial partner.

@@

The dataset used in this section was obtained from our partner and contains records representing delivery operations from up to 8 production centers over a period of 50 days. We generated 50 instances, with each instance corresponding to a daily operation. The instances include informations such as each driver capacity, associated depot, and shift start time. For each customer, we have the due time, total demand, and the number of orders (type of concrete) received. For each order, we specified its demand and associated production plant. Finally, we include loading capacities for each production center. 
The name of an instance has the format $C\_k\_n\_o\_z$. $k$ is the number of available drivers, $n$ the number of customers, $o$ the total number of daily orders, and $z$ the number of plants.

The dataset is subdivised in a small, medium and large sets, according to the number of drivers and total demands served. We summarized both sets in Table~\ref*{}
or 
The dataset is subdivised in a small, medium and large sets, according to the number of orders served. We summarized both sets in Table~\ref*{}

% This dataset allows us to evaluate the effectiveness and performance of our GRASP algorithm in optimizing delivery operations in a real-world setting.

@@

- resultats avec les instances (8 au total) où uniquement la flotte de Unibéton a éét utilisée. (utiliser les mêmes pour la comparaison avec les données réelles )
distance, demande servie, first, underwork, overtime.
- resultats avec les autres instances
- distance, demande servie, first delivery.

% df = df[df['Name'].isin([20201010,20201021,20201107,20201017,20201031,20201109,20201026,20201121])]

@@

\cite{kinable2014concrete} provided a benchmark dataset for their variant of the $CDP$. Although their primary objective is to maximize the total load delivered each day, without considering loading operations at a plant and ensuring deliveries within specified time windows, we can still use this dataset to analyze and compare the performance of our algorithm under different conditions. Table \ref*{tab:cdp_instances} shows the performance of the GRASP heuristic along with the other methods (CP, MIP, SD-heuristic, and Hybrid) used in \cite{kinable2014concrete}.  For each method, the table shows the average execution time in milliseconds and the corresponding gap between the obtained solution and the upper bound provided by \cite{kinable2014concrete}.  Overall, the GRASP algorithm shows promise in providing near-optimal solutions in a relatively short time for this benchmark of the CDP. It outperformed all methods except CP, achieving an average distance of 5.7\% (13.8\%) with a runtime of 13,186 (574,983) milliseconds for Set A (Set B). The SD-heuristic appears to be much faster, but GRASP provides a better balance between solution quality and computation time. It should also be noted that the exact MIP, CP, and Hybrid are initialized by the results of the SD-heuristic.

\begin{table}[htbp]
    \centering
    \caption{Performance of the GRASP heuristic on the CDP instances}
    \label{tab:cdp_instances}
    \resizebox{1\textwidth}{!}{
    \begin{tabular}{@{}cccccccccccc@{}}
    \toprule
     &  & \multicolumn{2}{c}{CP} & \multicolumn{2}{c}{MIP} & \multicolumn{2}{c}{SD-heuristic} & \multicolumn{2}{c}{Hybrid} & \multicolumn{2}{c}{GRASP} \\ \cmidrule(l){3-10} \cmidrule(l){11-12} 
     &  & Gap (\%) & Time (ms) & Gap (\%) & Time (ms) & Gap (\%) & Time (ms) & Gap (\%) & Time (ms) & Gap (\%) & Time (ms) \\ \cmidrule(l){3-10} \cmidrule(l){11-12} 
    \multirow{2}{*}{Set A} & Avg & 4.2 & 197,012 & 7.3 & 13,405,798 & 9.1 & 23 & 7.0 & 100,765 & 5.7 & 13,186 \\ \cmidrule(l){2-10} \cmidrule(l){11-12} 
     & Opt & \multicolumn{2}{c}{40/64} & \multicolumn{2}{c}{37/64} & \multicolumn{2}{c}{30/64} & \multicolumn{2}{c}{35/64} & \multicolumn{2}{c}{35/64} \\ \cmidrule(l){3-10} \cmidrule(l){11-12} 
     \multirow{2}{*}{Set B} & Avg & 12.1 & 357,313 & - & - & 16.3 & 1,331 & - & - & 13.8 & 574,983\\ \cmidrule(l){2-10} \cmidrule(l){11-12} 
     & Opt & \multicolumn{2}{c}{55/128} & \multicolumn{2}{c}{-} & \multicolumn{2}{c}{40/128} & \multicolumn{2}{c}{-} & \multicolumn{2}{c}{44/128} \\ \bottomrule
    \end{tabular}
    }
    \end{table}
